### 2021-04-12 doc/VerticalArrangementsSheuTaragin.tex
wrote up table 1 and figure 1

\documentclass[12pt]{article}
\usepackage{chicago}    % bibliography package
\usepackage{graphicx}   % insert PostScript figures
\usepackage{setspace}   % controls line spacing
\usepackage{amsmath,amsthm,amssymb,amstext} % controls equation entry and symbols
\usepackage{rotating}   % rotates graphics
\usepackage{soul}       % controls hyphenation
\usepackage{epsfig}     % helps with including graphics
\usepackage{pdflscape}  % helps with displaying rotated graphics in PDFs
\usepackage{lscape}     % helps with rotating pages
\usepackage{setspace}   % controls line spacing
\usepackage{caption}    % controls  captions
\usepackage{adjustbox}  % shrink tables
\usepackage[margin=1in]{geometry} % margins
\usepackage{hyperref}   % displays URLs

@@

For each combination of number of retailers, number of wholesalers, and bargaining parameter, we draw 1,000 different sets of market primitives.  This results in XX million merger simulations.  We then eliminate mergers where the merger is unprofitable to the merging firms, as well as markets that do not pass the Hypothetical Monopolist Test, yielding XX million markets.\footnote{The Hypothetical Monopolist Test requires that were a monopolist to jointly own all products in a candidate market, that firm would raise the price of at least one of the merging producers' products by at least a ``small but significant non-transitory increase in price'' (SSNIP), which we take to be 5\%. } All XX million markets treat as primitives the number of retailers, the number of wholesalers, the bargaining parameter, and the wholesaler and retailer marginal costs, which we allow to be either constant or linear.

@@

Table \ref{tab:simsum} provides summary statistics across our various simulations.\footnote{The \href{https://CRAN.R-project.org/package=antitrust}{\texttt{antitrust} R package} contains the computer code needed to calibrate and simulate the effects of mergers in a range of competitive scenarios, including the ones described here.}  The median average wholesale pre-merger price is almost \$5, and the median average retail pre-merger price is \$12.  Because the market size is set to 1, these average prices are equal to total pre-merger expenditures.  Pre-merger HHIs range between 2,029 at the $25^{th}$ percentile to 5,143 at the $75^{th}$, with a median of 2,828. HHIs for horizontal downstream mergers increase by 1,658 points at the median, resulting in a median post-merger HHI equal to 4,270. HHIs for upstream mergers increase by 1,179 points at the median, resulting in a median post-merger HHI equal to 4,243. HHIs for vertical mergers increase by 953 points at the median, resulting in a median post-merger HHI equal to 5,174.\footnote{We compute the post-merger HHI for vertical mergers by calculating the merged firms' market share as the sum of all the shares of downstream products that either incorporate the upstream partner's input or are sold by the downstream partner.  }  Many of these markets fall into the span designated by the DOJ/FTC Horizontal Merger Guidelines as ``Highly Concentrated Markets,'' with HHIs over 2,500.\footnote{HHI thresholds are discussed in the 2010 Horizontal Merger Guidelines, Section 5.3.  }

@@

We focus first on the results for consumers in the left-most panel of Figure \ref{fig:surplussum}.  The median change is negative, indicating harm, across all four types of mergers for both cost specifications.  However, the distributions and magnitudes differ.  In Bertrand markets, there is only a partial rank-ordering of consumer harm across different types of mergers: consumer harm from integrated mergers first-order stochastically dominates consumer harm from all other merger  types, while downstream and upstream mergers each first-order stochastically dominate consumer harm from vertical mergers, but do not stochastically dominate one other.   Median harm from downstream  mergers is 4.5\% of pre-merger total expenditures, 1.3 times the magnitude of that from upstream mergers, and 3.1 times the magnitude of that from vertical mergers. By contrast, under second score, consumer harm in upstream mergers stochastically dominates consumer harm in downstream mergers, which in turn dominates harm in vertical mergers. The median harm from upstream mergers is 3.5\% of pre-merger total expenditures, 4.1 times that of downstream mergers, and 15 times that of vertical mergers.

Four types of mergers are almost always harmful: upstream under both constant and linear marginal costs, vertical mergers with linear marginal costs, and integrated mergers with linear marginal costs.  That upstream horizontal mergers are net harmful is somewhat surprising, as it indicates that any EDM from the unintegrated upstream firm merging with its integrated rival is outweighed by the recapture effect.  Likewise, the fact that vertical mergers with linear marginal costs are almost never beneficial when marginal costs are linear is surprising, as it suggests that the incentive to raise rivals' cost dominates the benefit from EDM. %especially given that about 40\% of vertical mergers with constant marginal costs are beneficial.
\textbf{Is it worth exploring which effect dominates: the lessening of EDM or the strengthening of RRC? Maybe do this by allowing assymetric cost structures between merging and non-merging parties?}

@@

Harm is less prevalent for vertical mergers and downstream mergers with constant marginal costs.  Vertical mergers in Bertrand markets benefit consumers in about 40\% of all simulations , and downstream horizontal mergers benefit consumers in XX\% of all simulations.  The median consumer benefit among positive outcomes from vertical mergers equals 4.2\% of pre-merger total expenditures (comparable to the median consumer harm from downstream mergers under Bertrand), and the median consumer benefit from downstream mergers in auction markets equals 1.7\% of pre-merger revenues (less than two-fifths of the median consumer harm from downstream mergers under Bertrand).

Turning to retailers in the second panel of Figure \ref{fig:surplussum}, we find that while downstream mergers, vertical mergers and integrated mergers benefit retailers, upstream mergers frequently -- though not universally-- harm retailers. We also find that for all merger types, the median retailer surplus under constant marginal cost first-order stochastically dominate the retail surplus distribution under linear costs, indicating that the benefits from EDM are typically greater under constant marginal costs than linear marginal costs.

@@

As for wholesaler surplus, which appears in the third panel, the effects seen there are the mirror image of those for retailers, with net harm in almost all downstream mergers and the majority of vertical mergers, and net benefits in all upstream mergers.  Although one would expect wholesalers to be harmed by downstream mergers, the amount of harm in Bertrand markets is about one-third  of that in second score markets. This difference partially explains the difference in the magnitude of consumer harm between downstream mergers with Bertrand versus auctions: namely, retailers' greater ability to extract wholesaler surplus in second score auction markets translates into more savings passed thru to consumers, which offsets the harm from less retail competition.

@@

In terms of total welfare, with the exception of second score downstream mergers, our simulated mergers are typically net harmful, with second score vertical mergers yielding the highest median total harm (-2.5\% of pre-merger expenditures), followed by Bertrand downstream mergers (-2.2\%), Bertrand vertical mergers (-1.5\%), and then upstream mergers.  Downstream mergers under auctions are slightly beneficial at the median, with a gain of 0.5\% of pre-merger expenditures. About 16\% of Bertrand vertical mergers and 7\%  of second score vertical mergers are net beneficial.

@@

\begin{sidewaysfigure}
\centering
\includegraphics[scale=0.85]{../output/CVvertincumbBW.png}
\caption{The figure displays box and whisker plots summarizing the extent to which mergers affect consumer, retailer, wholesaler, and total surplus as the number of vertically integrrated firms present in a market change. Each blue box depicts the effects assuming that retailers are playing a Bertrand pricing game. Whiskers depict the $5^{th}$ and $95^{th}$ percentiles of a particular outcome, boxes depict the $25^{th}$ and $75^{th}$ percentiles, and the solid horizontal line depicts the median. }
\label{fig:CVvertincumbBW}
\end{sidewaysfigure}

>>>>

\documentclass[12pt]{article}
\usepackage{chicago}    % bibliography package
\usepackage{graphicx}   % insert PostScript figures
\usepackage{setspace}   % controls line spacing
\usepackage{amsmath,amsthm,amssymb,amstext} % controls equation entry and symbols
\usepackage{rotating}   % rotates graphics
\usepackage{soul}       % controls hyphenation
\usepackage{epsfig}     % helps with including graphics
\usepackage{pdflscape}  % helps with displaying rotated graphics in PDFs
\usepackage{lscape}     % helps with rotating pages
\usepackage{setspace}   % controls line spacing
\usepackage{caption}    % controls  captions
\usepackage{adjustbox}  % shrink tables
\usepackage[margin=1in]{geometry} % margins
\usepackage{hyperref}   % displays URLs
%packages for kable tables
 \usepackage{booktabs}
 \usepackage{longtable}
 \usepackage{array}
 \usepackage{multirow}
 \usepackage{wrapfig}
 \usepackage{float}
 \usepackage{colortbl}
 \usepackage{pdflscape}
 \usepackage{tabu}
 \usepackage{threeparttable}
 \usepackage{threeparttablex}
 \usepackage[normalem]{ulem}
 \usepackage{makecell}
 \usepackage{xcolor}

@@

For each combination of number of retailers, number of wholesalers, and bargaining parameter, we draw 1,000 different sets of market primitives.  This results in 2.2 million merger simulations.  We then eliminate mergers where the merger is unprofitable to the merging firms, as well as markets that do not pass the Hypothetical Monopolist Test, yielding XX million markets.\footnote{The Hypothetical Monopolist Test requires that were a monopolist to jointly own all products in a candidate market, that firm would raise the price of at least one of the merging producers' products by at least a ``small but significant non-transitory increase in price'' (SSNIP), which we take to be 5\%. } All 2.2 million markets treat as primitives the number of retailers, the number of wholesalers, the bargaining parameter, and the wholesaler and retailer marginal costs, which we allow to be either constant or linear.

@@

Table \ref{tab:simsum} provides summary statistics across our various simulations.\footnote{The \href{https://CRAN.R-project.org/package=antitrust}{\texttt{antitrust} R package} contains the computer code needed to calibrate and simulate the effects of mergers in a range of competitive scenarios, including the ones described here.}  The median average wholesale pre-merger price is almost \$5, and the median average retail pre-merger price is \$13.  Because the market size is set to 1, these average prices are equal to total pre-merger expenditures.  Pre-merger HHIs range between 2,629 at the $25^{th}$ percentile to 4,287 at the $75^{th}$, with a median of 3,688. HHIs for horizontal downstream mergers increase by 2,027 points at the median, resulting in a median post-merger HHI equal to 4,963. HHIs for upstream mergers increase by 2,040 points at the median, resulting in a median post-merger HHI equal to 5,207. HHIs for vertical mergers increase by 1,292 points at the median, resulting in a median post-merger HHI equal to 6,016.\footnote{We compute the post-merger HHI for vertical mergers by calculating the merged firms' market share as the sum of all the shares of downstream products that either incorporate the upstream partner's input or are sold by the downstream partner.  }  HHIs for integrated mergers increase by 2,800 points at the median, resulting in a median post-merger HHI equal to 7,272. Many of these markets fall into the span designated by the DOJ/FTC Horizontal Merger Guidelines as ``Highly Concentrated Markets,'' with HHIs over 2,500.\footnote{HHI thresholds are discussed in the 2010 Horizontal Merger Guidelines, Section 5.3.  }

@@

We focus first on the results for consumers in the left-most panel of Figure \ref{fig:surplussum}.  The median change is negative, indicating harm, across all four types of mergers for both cost specifications.  However, the distributions and magnitudes differ.  In Bertrand markets, there is only a partial rank-ordering of consumer harm across different types of mergers: consumer harm from integrated mergers first-order stochastically dominates consumer harm from all other merger  types, while downstream and upstream mergers each first-order stochastically dominate consumer harm from vertical mergers, but do not stochastically dominate one other.   Median harm from integrated mergers is 20\% of pre-merger total expenditures, more than 2 times the magnitude of that from downstream mergers, 3.1 times the magnitude of that from upstream mergers, and 5.4 times the magnitude of vertical mergers. Consumer harm under linear marginal costs first-order stochastically dominates consumer harm under constant marginal costs. The median harm from markets with linear marginal costs is about 6.8\% of pre-merger total expenditures, 1.2 times the magnitude of markets with constant marginal costs.

Four types of mergers are unlikely to benefit consumers: upstream under both constant and linear marginal costs, vertical mergers with linear marginal costs, and integrated mergers with linear marginal costs.  That upstream horizontal mergers are largely net harmful is somewhat surprising, as it indicates that any EDM from the unintegrated upstream firm merging with its integrated rival is outweighed by the recapture effect. Similar logic applies to integrated mergers. Likewise, the fact that vertical mergers with linear marginal costs are almost never beneficial when marginal costs are linear is surprising, as it suggests that the incentive to raise rivals' cost dominates the benefit from EDM. %especially given that about 40\% of vertical mergers with constant marginal costs are beneficial.
\textbf{Is it worth exploring which effect dominates: the lessening of EDM or the strengthening of RRC? Maybe do this by allowing assymetric cost structures between merging and non-merging parties?}

@@

Harm is less prevalent for vertical mergers, downstream mergers and integrated mergers when firms have constant marginal costs.  Vertical mergers when firms have constant marginal costs benefit consumers in about 33\% of all simulations , downstream horizontal mergers benefit consumers in about 13\% of all simulations, and integrated mergers benefit consumers in about 4.9\% of all simulations.  The median consumer benefit among positive outcomes equals 5.9\% of pre-merger total expenditures from vertical mergers, 9.5\% of pre-merger total expenditures from downstream mergers, and 4.9\% of integrated mergers.

Turning to retailers in the second panel of Figure \ref{fig:surplussum}, we find that while downstream mergers, vertical mergers and integrated mergers always benefit retailers, upstream mergers  harm retailers in about 79\% of all simulation. Moreover, there is a partial rank-ordering across mergers: the retailer surplus distribution from integrated mergers first-order stochastically dominates the retailer surplus distribution from downstream mergers, which dominates the retailer surplus distribution from upstream mergers, but not the retailer surplus distribution from vertical mergers.  We also find that  for vertical and integrated mergers, retailer surplus under constant marginal cost first-order stochastically dominates the retail surplus distribution under linear costs, again suggesting that the incentive to raise rivals' costs dominates the benefits from EDM. By contrast, for downstream and upstream mergers, there is no clear rank-ordering between markets with constant marginal costs and markets with linear costs.

@@

As for wholesaler surplus, which appears in the third panel, the effects seen there are largely the mirror image of those for retailers: upstream mergers increase wholesaler surplus in about 97\% of all simulated mergers, while the other merger types decrease wholesaler surplus from anywhere between 81\% (downstream mergers) and 68\% (integrated mergers) of all simulated mergers. \textbf{comparing these results to Sheu/Taragin, the effects are less stark (ie less of a transfer.} Again, we see a a partial rank-ordering across mergers: the retailer surplus distribution from upstream mergers first-order stochastically dominates the retailer surplus distribution from downstream mergers, which dominates the retailer surplus distribution from upstream mergers, but not the retailer surplus distribution from vertical mergers.

@@

In terms of total welfare, with the exception of the approximately 16\% of vertical mergers with constant costs that are beneficial, our simulated mergers are almost always net harmful. Moreover there is a complete rank ordering of mergers, with total harm from integrated mergers first-order stochastically dominating total harm from downstream mergers, which dominates total harm from upstream mergers, which dominates totals harm from vertical mergers. Finally, there is little difference in total harm under constant and linear costs, though total harm under linear costs continues to first-order stochastically dominate total harm under constant marginal costs for vertical and integrated mergers.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Tables and Figures
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}
\include{sumtable}
\label{tab:simsum}
\caption{Summary Statistics}
\end{table}

@@

\begin{sidewaysfigure}
\centering
\includegraphics[scale=0.85]{../output/CVvertincumbBW.png}
\caption{The figure displays box and whisker plots summarizing the extent to which mergers affect consumer, retailer, wholesaler, and total surplus as the number of vertically integrated firms present in a market change. Each blue box depicts the effects assuming that retailers are playing a Bertrand pricing game. Whiskers depict the $5^{th}$ and $95^{th}$ percentiles of a particular outcome, boxes depict the $25^{th}$ and $75^{th}$ percentiles, and the solid horizontal line depicts the median. }
\label{fig:CVvertincumbBW}
\end{sidewaysfigure}
